---
title: "Fake News Data - Correlation"
author: "Natália Tosi"
date: "08/18/2021"
output: pdf_document
---

```{r message = FALSE, warning = FALSE, include = FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 8, 
  fig.height = 4
  )

options(scipen = 999)
rm(list=ls())

library(tidyverse)
library(broom) 
library(stargazer)
library(knitr)
library(car)
library(Hmisc)
library(kableExtra)

```

\bigskip

```{r}

data_code <- read_csv("fake_news_data_code.csv") %>%
  select(-1)

```

\bigskip


# Model Variables


\bigskip

**Dependent Variable = Shared Fake News (`shared_fake_news_19`)** 

\bigskip

P19 - Have you ever shared a political news story online that you thought at the time was made up? (Single Answer)

1 = Yes; 0 = No

\bigskip

**Independent Variables**

\bigskip

1) Gender
2) Age
4) Income (class)
5) Education
6) Political Orientation
7) Religion (Evangelicals)

\bigskip

# Data

\bigskip

- P19 = shared_fake_news
- frequency_fake_news
- P21 = unnoticed_share
- public
- gov and politicians (create one single)
- social media

```{r}

data_fake_news <- data_code %>%
  select(sex, age_full, age_60, pol_orientation, P19, frequency_fake_news, P21,
         resp_population, resp_gov, resp_politicians, resp_press, 
         resp_social_media, severity_fake_news, shared_fake_news_19, education, 
         income, class, religion, evaluation, approval, shared_fake_news_19, 
         race_adj, sex_men, region_North, region_Northeast, `region_Center-West`,
         region_Southeast, region_South, religion_Catholic, religion_Evangelicals,
         `religion_Other religion`, `religion_No religion`, `age_60_16-24 age`, 
         `age_60_25-34 age`, `age_60_35-44 age`, `age_60_45-59 age`, 
          `age_60_60 or more`, evaluation_Unsure, P21_No, P21_Unsure, P21_Yes,
         capital_metrop, pol_orientation_right, pol_orientation_center,
         pol_orientation_left, pol_orientation_none, race_is_white,
         education_high, income_low, class_ab, class_c, class_de, has_religion) %>%
  rename(unnoticed_share = P21_Yes) %>%
  mutate(resp_gov_politicians = if_else(((resp_gov == 1) | (resp_politicians == 1)), 
                                        1, 0),
         sex_women = if_else(sex_men == 0, 1, 0))

```

\bigskip

# Correlation Matrix

\bigskip

```{r}

variables <- c("sex_men", "age_full", "class_ab", "class_c", "pol_orientation_right", 
               "pol_orientation_center", "pol_orientation_left", "education_high", 
               "religion_Evangelicals", "frequency_fake_news", "unnoticed_share",
               "resp_population", "resp_gov_politicians", "resp_social_media")
                              


variables_rename <- c("Gender(male)", "Age", "Class AB", "Class C", "Politics (R)", 
                      "Politics (C)", "Politics (L)", "Educated", "Evangelical",
                      "FN Freq", "Unnoticed FN", "Resp. Pop", "Resp. Gov", 
                      "Resp. SM")

```


```{r}

correlation_matrix <- function(df, 
                               type = "pearson",
                               digits = 3, 
                               decimal.mark = ".",
                               use = "all", 
                               show_significance = TRUE, 
                               replace_diagonal = FALSE, 
                               replacement = ""){
  
  # check arguments
  stopifnot({
    is.numeric(digits)
    digits >= 0
    use %in% c("all", "upper", "lower")
    is.logical(replace_diagonal)
    is.logical(show_significance)
    is.character(replacement)
  })
  # we need the Hmisc package for this
  require(Hmisc)
  
  # retain only numeric and boolean columns
  isNumericOrBoolean = vapply(df, function(x) is.numeric(x) | is.logical(x), logical(1))
  if (sum(!isNumericOrBoolean) > 0) {
    cat('Dropping non-numeric/-boolean column(s):', paste(names(isNumericOrBoolean)[!isNumericOrBoolean], collapse = ', '), '\n\n')
  }
  df = df[isNumericOrBoolean]
  
  # transform input data frame to matrix
  x <- as.matrix(df)
  
  # run correlation analysis using Hmisc package
  correlation_matrix <- Hmisc::rcorr(x, type = )
  R <- correlation_matrix$r # Matrix of correlation coeficients
  p <- correlation_matrix$P # Matrix of p-value 
  
  # transform correlations to specific character format
  Rformatted = formatC(R, format = 'f', digits = digits, decimal.mark = decimal.mark)
  
  # if there are any negative numbers, we want to put a space before the positives to align all
  if (sum(R < 0) > 0) {
    Rformatted = ifelse(R > 0, paste0(' ', Rformatted), Rformatted)
  }
  
  # add significance levels if desired
  if (show_significance) {
    # define notions for significance levels; spacing is important.
    stars <- ifelse(is.na(p), "   ", ifelse(p < .001, "***", ifelse(p < .01, "** ", ifelse(p < .05, "*  ", "   "))))
    Rformatted = paste0(Rformatted, stars)
  }
  # build a new matrix that includes the formatted correlations and their significance stars
  Rnew <- matrix(Rformatted, ncol = ncol(x))
  rownames(Rnew) <- colnames(x)
  colnames(Rnew) <- paste(colnames(x), "", sep =" ")
  
  # replace undesired values
  if (use == 'upper') {
    Rnew[lower.tri(Rnew, diag = replace_diagonal)] <- replacement
  } else if (use == 'lower') {
    Rnew[upper.tri(Rnew, diag = replace_diagonal)] <- replacement
  } else if (replace_diagonal) {
    diag(Rnew) <- replacement
  }
  
  return(Rnew)
}

```


```{r}
data_fake_news <- data_fake_news[, variables]

colnames(data_fake_news) <- variables_rename

correlation <- correlation_matrix(data_fake_news)

correlation_1 <- correlation[,c(1:7)]

correlation_2 <- correlation[,c(8:14)]
```


\newpage

```{r}

kable(correlation_1, caption = "Correlations between independent variables",
      booktabs = T, linesep = "") %>%
  kable_styling(latex_options = "striped", font_size = 8) %>% 
  row_spec(0, bold = TRUE)

```


```{r}

kable(correlation_2, caption = "Correlations between independent variables",
      booktabs = T, linesep = "") %>%
  kable_styling(latex_options = "striped", font_size = 8) %>% 
  row_spec(0, bold = TRUE)

```


