---
title: "Fake News Data - Demographics"
author: "Natália Tosi"
date: "08/10/2021"
output: pdf_document
---

```{r message = FALSE, warning = FALSE, include = FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 8, 
  fig.height = 4
  )

options(scipen = 999)

library(tidyverse)
library(broom) 
library(stargazer)
library(knitr)

```

\bigskip

```{r}

data_code <- read_csv("fake_news_data_code.csv") %>%
  select(-1)

```


# Demographics

\bigskip

**Sex**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(sex) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```
\bigskip

**Age**

\bigskip

```{r, echo = FALSE}

data_code %>%
   summarise(mean = mean(age_full),
             median = median(age_full),
             sd = sd(age_full)) %>%
  kable(digits = 3)

```

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(age_60) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```

\bigskip

**Race**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(race) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```

\bigskip

**Education**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(education) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```
\bigskip

**Class**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(class) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```
\bigskip

**Religion**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(religion) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```

\bigskip


**Region**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(region) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```
\bigskip

**City type**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(type) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```

\bigskip


**Political Orientation**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(pol_orientation) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```

\bigskip

**Government Approval Rating**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(approval) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```

\bigskip


# Answers Paper Demographics

**P19 - Have you ever shared a political news story online that you thought at the time was made up? (Single Answer)**

--> excluded "Unsure"

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(P19) %>%
  mutate(share = (n/sum(n))*100) %>%
  arrange(desc(P19)) %>%
  kable(digits = 3)

```


\bigskip

**P20 - How often do you come across news stories online that you think are almost completely made up? (Single Answer)**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(frequency_fake_news) %>%
  mutate(share = (n/sum(n))*100) %>%
  kable(digits = 3)

```


\bigskip

**P21 - Have you ever shared a political news story online that you later found out was made up? (Single Answer)**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(P21) %>%
  mutate(share = (n/sum(n))*100) %>%
  mutate(P21 = factor(P21, levels = c("Yes", "No", "Unsure"))) %>%
  arrange(P21) %>%
  kable(digits = 3)

```


\bigskip

**P22 - [IS ANSWERED YES] What was your reaction when you found out that the information shared was not true? If it happened more than once, select the option that was most frequent (Single Answer)**

\bigskip

```{r, echo = FALSE}

data_code %>%
  count(reaction_fake_news) %>%
  filter(!is.na(reaction_fake_news)) %>%
  mutate(share = (n/sum(n))*100) %>%
  arrange(desc(share)) %>%
  kable(digits = 3)

```


\bigskip

**P23 - How much responsibility does each of the following have in trying to prevent made up (fake news) stories from gaining attention”. (Single Answer per category)**

\bigskip

1 = "A great deal of responsibility" and "A fair amount of responsibility"

\bigskip

```{r}

resp_population <- data_code %>%
  count(resp_population) %>%
  mutate(population = (n/sum(n))*100) %>%
  select(-n)

resp_gov <- data_code %>%
  count(resp_gov) %>%
  mutate(governments = (n/sum(n))*100) %>%
  select(-n)

resp_politicians <- data_code %>%
  count(resp_politicians) %>%
  mutate(politicians = (n/sum(n))*100) %>%
  select(-n)

resp_press <- data_code %>%
  count(resp_press) %>%
  mutate(press = (n/sum(n))*100) %>%
  select(-n)

resp_social_media <- data_code %>%
  count(resp_social_media) %>%
  mutate(social_media = (n/sum(n))*100) %>%
  select(-n)

responsibility <- resp_population %>%
  left_join(resp_gov, by = c("resp_population" = "resp_gov")) %>%
  left_join(resp_politicians, by = c("resp_population" = "resp_politicians")) %>%
  left_join(resp_press, by = c("resp_population" = "resp_press")) %>%
  left_join(resp_social_media, by = c("resp_population" = "resp_social_media")) %>%
  mutate(resp_population = factor(resp_population, levels = c(1, 0))) %>%
  arrange(resp_population) %>%
  rename(c(answer = resp_population))
  
kable(responsibility,
      digits = 3) 

```

