---
title: "Fake News Data - Fake News Paper - Demographics (Only used in the paper + Evangelicals)"
author: "Natália Tosi"
date: "08/18/2021"
output: pdf_document
---

```{r message = FALSE, warning = FALSE, include = FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 8, 
  fig.height = 4
  )

options(scipen = 999)
rm(list = ls())

library(tidyverse)
library(broom) 
library(stargazer)
library(knitr)
library(car)
library(DescTools)


```

\bigskip

```{r}

data_code <- read_csv("fake_news_data_code.csv") %>%
  select(-1)

```

\bigskip

# Paper Research Questions (ADJUSTED)

\bigskip

  - RQ1. How do demographics (age, gender, religion, education, political orientation and income), affect the probability to share political fake news online?
  
  - RQ2. What is the interaction between political orientation and gender?
  
  - RQ3. How does the perception of frequency of political fake news online affect the
probability to share them?

  - RQ4. How does previous online fake news sharing (unnoticed), affect the probability
to share political fake news online (on purpose)?

  - RQ5. How does the perception of responsibility of 1) members of the public, 2)
government, politicians and elected officials and 3) Facebook and Twitter, in trying to prevent fake news affect the probability to share them online?

\bigskip

# Model Variables

\bigskip

**Dependent Variable = Shared Fake News (`shared_fake_news_19`)** 

\bigskip

P19 - Have you ever shared a political news story online that you thought at the time was made up? (Single Answer)

1 = Yes; 0 = No

\bigskip

**Independent Variables**

\bigskip

1) Gender
2) Age
4) Income (class)
5) Education
6) Political Orientation
7) Religion (Evangelical)


\bigskip


# Data

\bigskip

- P19 = shared_fake_news
- frequency_fake_news
- P21 = unnoticed_share
- public
- gov and politicians (create one single)
- social media

```{r}

data_fake_news <- data_code %>%
  select(sex, age_full, age_60, pol_orientation, P19, frequency_fake_news, P21,
         resp_population, resp_gov, resp_politicians, resp_press, 
         resp_social_media, severity_fake_news, shared_fake_news_19, education, 
         income, class, religion, evaluation, approval, shared_fake_news_19, 
         race_adj, sex_men, region_North, region_Northeast, `region_Center-West`,
         region_Southeast, region_South, religion_Catholic, religion_Evangelicals,
         `religion_Other religion`, `religion_No religion`, `age_60_16-24 age`, 
         `age_60_25-34 age`, `age_60_35-44 age`, `age_60_45-59 age`, 
          `age_60_60 or more`, evaluation_Unsure, P21_No, P21_Unsure, P21_Yes,
         capital_metrop, pol_orientation_right, pol_orientation_center,
         pol_orientation_left, pol_orientation_none, race_is_white,
         education_high, income_low, class_ab, class_c, class_de, has_religion) %>%
  rename(unnoticed_share = P21_Yes) %>%
  mutate(resp_gov_politicians = if_else(((resp_gov == 1) | (resp_politicians == 1)), 
                                        1, 0),
         sex_women = if_else(sex_men == 0, 1, 0))

```


\bigskip

# Regressions - All categories


\bigskip

## Regression 1 - Share fake news with demographics (sex, age, income, education, political orientation)

\bigskip

```{r}

reg_1 <- glm(shared_fake_news_19 ~ sex_men + age_full + class_ab + 
               class_c + pol_orientation_right + pol_orientation_center +
               pol_orientation_left + education_high + religion_Evangelicals, 
               family = binomial(link = 'logit'),
               data = data_fake_news)

#Left out: women, class DE, no political orientation

```

\bigskip

## Regression 2 - How does the perception of frequency of political fake news online affect the probability to share them?

\bigskip

```{r}

reg_2 <- glm(shared_fake_news_19 ~ frequency_fake_news + sex_men + age_full + class_ab + 
               class_c + pol_orientation_right + pol_orientation_center +
               pol_orientation_left + education_high + religion_Evangelicals,
               family = binomial(link = 'logit'),
               data = data_fake_news)

```

\bigskip

## Regression 3 - How does previous online fake news sharing (unnoticed), affect the probability to share political fake news online (on purpose)?

\bigskip

```{r}

reg_3 <- glm(shared_fake_news_19 ~ unnoticed_share + frequency_fake_news + 
               sex_men + age_full + class_ab + class_c + pol_orientation_right + 
               pol_orientation_center + pol_orientation_left + education_high +
               religion_Evangelicals,
               family = binomial(link = 'logit'),
               data = data_fake_news)

```

\bigskip

## Regression 4 - How does the perception of responsibility of 1) members of the public, 2) government, politicians and elected officials and 3) Facebook and Twitter, in trying to prevent fake news affect the probability to share them online?

\bigskip

```{r}

reg_4 <- glm(shared_fake_news_19 ~ resp_population + resp_gov_politicians + 
               resp_social_media + unnoticed_share + frequency_fake_news + 
               sex_men + age_full + class_ab + class_c + pol_orientation_right + 
               pol_orientation_center + pol_orientation_left + education_high +
               religion_Evangelicals,
               family = binomial(link = 'logit'),
               data = data_fake_news)

```


\bigskip

## Regression 5 - What is the interaction between political orientation and gender?

\bigskip

```{r}

reg_5 <- glm(shared_fake_news_19 ~ resp_population + resp_gov_politicians + 
               resp_social_media + unnoticed_share + frequency_fake_news + 
               sex_men + age_full + class_ab + class_c + pol_orientation_right + 
               pol_orientation_center + pol_orientation_left + education_high + 
               sex_men*pol_orientation_right + sex_men*pol_orientation_center + 
               sex_men*pol_orientation_left + religion_Evangelicals,
               family = binomial(link = 'logit'),
               data = data_fake_news)

```

\bigskip

```{r, results = 'asis', message = FALSE, echo = FALSE}

stargazer(reg_1, reg_2, reg_3, reg_4, reg_5,
          title = "Likelihood of Sharing Fake News - Logit Models (Estimates)", 
          type = "latex", 
          digits = 3, 
          no.space = TRUE, 
          model.numbers = FALSE,
          dep.var.labels = c("Shared FN"),
          column.labels = c("Model 1", "Model 2", "Model 3", "Model 4", "Model 5"),
          covariate.labels = c("Resp. Pop", "Resp. Gov", "Resp. SM", "Unnoticed FN",
                               "FN Freq", "Gender(male)", "Age", "Class AB", "Class C",
                               "Politics (R)", "Politics (C)", "Politics (L)", "Educated",
                               "Evangelical", "Gender(male)*Pol(R)", "Gender(male)*Pol(C)",
                               "Gender(male)*Pol(L)"),
          header = FALSE,
          font.size = "small",
          column.sep.width = "-15pt",
          add.lines = list(c("Nagelkerke R$^2$", "0.021", "0.036", "0.247", "0.248","0.252"),
                           c("Cox and Snell R$^2$", "0.013", "0.023", "0.159", "0.16", "0.162")))

```


```{r, eval = FALSE, echo = FALSE}

cox_snell_R2 <- as.tibble(t(c("reg_1", "reg_2", "reg_3", "reg_4", "reg_5"))) 
colnames(cox_snell_R2) <- cox_snell_R2[1,]

cox_snell_R2[[1]] <- round(PseudoR2(reg_1, which = "CoxSnell")[[1]],3)
cox_snell_R2[[2]] <- round(PseudoR2(reg_2, which = "CoxSnell")[[1]],3)
cox_snell_R2[[3]] <- round(PseudoR2(reg_3, which = "CoxSnell")[[1]],3)
cox_snell_R2[[4]] <- round(PseudoR2(reg_4, which = "CoxSnell")[[1]],3)
cox_snell_R2[[5]] <- round(PseudoR2(reg_5, which = "CoxSnell")[[1]],3)

cox_snell_R2

```

```{r, eval = FALSE, echo = FALSE}

nagelkerke_R2 <- as.tibble(t(c("reg_1", "reg_2", "reg_3", "reg_4", "reg_5"))) 
colnames(nagelkerke_R2) <- nagelkerke_R2[1,]

nagelkerke_R2[[1]] <- round(PseudoR2(reg_1, which = "Nagelkerke")[[1]],3)
nagelkerke_R2[[2]] <- round(PseudoR2(reg_2, which = "Nagelkerke")[[1]],3)
nagelkerke_R2[[3]] <- round(PseudoR2(reg_3, which = "Nagelkerke")[[1]],3)
nagelkerke_R2[[4]] <- round(PseudoR2(reg_4, which = "Nagelkerke")[[1]],3)
nagelkerke_R2[[5]] <- round(PseudoR2(reg_5, which = "Nagelkerke")[[1]],3)

nagelkerke_R2

```



```{r}

plogis(coef(reg_1))


odds <- exp(coef(reg_1))

probability <-  odds / (1 + odds)
probability


```


\newpage

# Variance Inflation Factor (VIF)

\bigskip

"For a given predictor (p), multicollinearity can assessed by computing a score called the variance inflation factor (or VIF), which measures how much the variance of a regression coefficient is inflated due to multicollinearity in the model.

The smallest possible value of VIF is one (absence of multicollinearity). As a rule of thumb, a VIF value that exceeds 5 or 10 indicates a problematic amount of collinearity (James et al. 2014)."

```{r}

model_1 <- as.data.frame(car::vif(reg_1)) %>%
  rownames_to_column("variables")

model_2 <- as.data.frame(car::vif(reg_2)) %>%
  rownames_to_column("variables")

model_3 <- as.data.frame(car::vif(reg_3)) %>%
  rownames_to_column("variables")

model_4 <- as.data.frame(car::vif(reg_4)) %>%
  rownames_to_column("variables")

model_5 <- as.data.frame(car::vif(reg_5)) %>%
  rownames_to_column("variables")

vif_test <- model_1 %>%
  left_join(model_2, by = c("variables" = "variables")) %>%
  left_join(model_3, by = c("variables" = "variables")) %>%
  left_join(model_4, by = c("variables" = "variables")) %>%
  left_join(model_5, by = c("variables" = "variables"))

names(vif_test) <- c("variables", "model_1", "model_2", "model_3",
                     "model_4", "model_5")

vif_test %>%
  kable(caption = "Variance Inflation Factor (VIF) per variable and model",
      align = "c")

```
