---
title: "Fake_News_Data_Cleaning"
author: "Natália Tosi"
date: "8/3/2021"
output: pdf_document
---

```{r message = FALSE, warning = FALSE, include = FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 8, 
  fig.height = 4
  )

options(scipen = 999)

library(tidyverse)
library(fastDummies)
library(broom) 
library(stargazer)

```

\bigskip

# Importing Data and Labels

\bigskip

```{r}

data_raw_port <- read_csv(
  "BANCO_NACIONAL_FAKENEWS_2021-08-03_CLEAN - label.csv")

variable_names <- read_csv(
  "BANCO_NACIONAL_FAKENEWS_2021-08-03_CLEAN - variable_names.csv")

answers_labels <- read_csv(
  "BANCO_NACIONAL_FAKENEWS_2021-08-03_CLEAN - answers_translated.csv")

```

\bigskip

# Translating Document

\bigskip

```{r}

data_clean <- as_tibble(data_raw_port)
colnames(data_clean) <- as_vector(variable_names[,2])

data_eng <- data_clean
data_eng[-c(1,2,3,7)] <- lapply(data_clean[-c(1,2,3,7)], 
      function(x) answers_labels$answers_eng[match(x, 
                                        answers_labels$answers_port)])

```

\bigskip

Save as new CSV file

\bigskip

```{r}

write.csv(data_eng,'fake_news_db_english.csv')

```

\bigskip

# Summary Statistics

\bigskip

```{r}

data_eng <- data_eng %>%
  mutate(
    evaluation = case_when(
    P1 %in% c("Excellent", "Good") ~ "Excellent/Good",
    P1 %in% c("Bad", "Terrible") ~ "Bad/Terrible",
    TRUE ~ P1),
    approval = case_when(
      P2 %in% c("Strongly approves", "Approves") ~ "Approves",
      P2 %in% c("Strongly disapproves", "Disapproves") ~ "Disapproves",
      TRUE ~ P2))


data_fake_news_dem <- data_eng %>%
  select(idInterview, state, region, type, sex, age_full, age_60, evaluation, 
         approval, P4, P19, P20, P21, P23_1, P23_2, P23_3, P23_4, P23_5, 
         education_full, race, religion_full, income_full, class_full, age_50, 
         education, income, class, religion) %>%
  mutate(shared_fake_news = if_else(P19 == "Yes", 1, 0))


data_fake_news_dem <- data_fake_news_dem %>%
  mutate(sex = factor(sex, levels = c("Men", "Women")),
         region = factor(region, levels = c("North", "Northeast", "Center-West", 
                                   "Southeast", "South")),
         type = factor(type, levels = c("Capital", "Metropolitan region", 
                                        "Countryside")),
         evaluation = factor(evaluation, levels = c("Excellent/Good", "Regular", 
                                        "Bad/Terrible", "Unsure")),
         approval = factor(approval, levels = c("Approves", 
                                     "Neither approves nor disapproves",
                                      "Disapproves", "Unsure")),
         P4 = factor(P4, levels = c("Right/Center-Right", "Center", 
                                    "Left/Center-Left", 
                            "I no longer have a defined political orientation",
                            "I never had a political orientation", "Unsure")),
         race = factor(race, levels = c("White", "Black", "Pardo (brown)", 
                                        "Indigenous", "Yellow", "Other")),
         education = factor(education, levels = c("No education", "Elementary School", 
                                       "High School", "Higher Education")),
         income = factor(income, levels = c("Up to 1 MW", "1 to 3 MWs", "3 to 6 MWs", 
                                    "More than 6 MWs", "Did not answer")),
         class = factor(class, levels = c("A/B", "C", "D/E", "DN/DA")),
         religion = factor(religion, levels = c("Catholic", "Evangelicals", 
                                      "Other religion", "No religion")))


#data_fake_news_dem <- data_fake_news_dem %>%
 # mutate_if(is.character, as.factor) %>%
  #dummy_cols(select_columns = c("region", "type", "sex", "evaluation", 
   #      "approval", "P4", "P19", "P20", "P21", "P23_1", "P23_2", "P23_3", 
    #     "P23_4", "P23_5", "race", "education", "income", "class", "religion"))


```

#DEMOGRAPHICS

Sex

```{r}

data_fake_news_dem %>%
  count(sex) %>%
  mutate(share = n/sum(n))

```

Region

```{r}

data_fake_news_dem %>%
  count(region) %>%
  mutate(share = n/sum(n))

```

City type

```{r}

data_fake_news_dem %>%
  count(type) %>%
  mutate(share = n/sum(n))


```
Age

```{r}

data_fake_news_dem %>%
   summarise(mean = mean(age_full),
             median = median(age_full),
             sd = sd(age_full))

```

Political Orientation

```{r}

data_fake_news_dem <- rename(data_fake_news_dem, pol_orientation = P4)

data_fake_news_dem %>%
  count(pol_orientation) %>%
  mutate(share = n/sum(n))

```

Government Approval Rating

```{r}

data_fake_news_dem %>%
  count(approval) %>%
  mutate(share = n/sum(n))

```


Race

```{r}

data_fake_news_dem %>%
  count(race) %>%
  mutate(share = n/sum(n))

```


Education

```{r}

data_fake_news_dem %>%
  count(education) %>%
  mutate(share = n/sum(n))

```

Class

```{r}

data_fake_news_dem %>%
  count(class) %>%
  mutate(share = n/sum(n))

```

Religion

```{r}

data_fake_news_dem %>%
  count(religion) %>%
  mutate(share = n/sum(n))

```


# DUMMIES

  sex_men: 1 Men, 0 Women;
  region: 5 levels;
  capital_metrop: 1 Capital and Metropolitan region, 0 Countryside;
  approvaes_gov: 1 Approves, 0 Neither approves nor disapproves, Disapproves, Unsure;
  pol_orientation: Right/Center-Right, Center, Left/Center-Left, No orientation 
        (I no longer have a defined political orientation, I never had a political
        orientation, Unsure);
  race_is_white: 1 White, 0 Black, Pardo (brown), Indigenous, Yellow, Other;
  education_high: 1 High School and Higher Education, 0 No education and Elementary School, 
  income_low: 1 Up to 1 MW, 1 to 3 MWs, and Did not answer, 0 3 to 6 MWs and 
        More than 6 MWs;
  class: 3 levels: A/B, C, D/E and DN/DA;
  religion: 4 levels: Catholic, Evangelicals, Other religion, No religion


```{r}

data_fake_news_dem <- data_fake_news_dem %>%
  mutate(sex_men = if_else(sex == "Men", 1, 0)) %>%
  dummy_cols(select_columns = c("region")) %>%
  mutate(capital_metrop = if_else(type %in% c("Capital", "Metropolitan region"), 1, 0),
         approves_gov = if_else(approval == "Approves", 1, 0),
         pol_orientation_right = if_else(pol_orientation == "Right/Center-Right", 1, 0),
         pol_orientation_center = if_else(pol_orientation == "Center", 1, 0),
         pol_orientation_left = if_else(pol_orientation == "Left/Center-Left", 1, 0),
         pol_orientation_none = if_else(pol_orientation %in% c(
           "I no longer have a defined   political orientation", 
           "I never had a political orientation", "Unsure"), 1, 0),
         race_is_white = if_else(race == "White", 1, 0),
         education_high = if_else(education %in% c("High School", "Higher Education"), 1, 0),
         income_low = if_else(income %in% c("Up to 1 MW", "1 to 3 MWs", "Did not answer"),
                              1, 0),
         class_ab = if_else(class == "A/B", 1, 0),
         class_c = if_else(class == "C", 1, 0),
         class_de = if_else(class %in% c("D/E", "DN/DA"), 1, 0)) %>%
  dummy_cols(select_columns = c("religion"))

```



# LIKELIHOOD OF SHARING FAKE NEWS


```{r}

data_fake_news_dem %>%
  count(shared_fake_news) %>%
  mutate(share = n/sum(n))

```


```{r}

sapply(data_fake_news_dem,function(x) sum(is.na(x)))
sapply(data_fake_news_dem, function(x) length(unique(x)))

```





Model 1 - Only demographics

```{r}

model_1 <- glm(shared_fake_news ~ sex_men + age_full + race_is_white + 
                education_high + income_low + class_c,
               family = binomial(link = 'logit'),
               data = data_fake_news_dem)

summary(model_1)

```

Model 2 - Demographics + Political Orientation

```{r}

model_2 <- glm(shared_fake_news ~ sex_men + age_full + race_is_white + 
                education_high + income_low + class_c + pol_orientation_right +
                pol_orientation_center + pol_orientation_left,
               family = binomial(link = 'logit'),
               data = data_fake_news_dem)

summary(model_2)

```


Model 3 - Demographics + Political Orientation + City and Region


```{r}

model_3 <- glm(shared_fake_news ~ sex_men + age_full + race_is_white + 
                education_high + income_low + class_c + pol_orientation_right +
                pol_orientation_center + pol_orientation_left + region_North +
                region_Northeast + `region_Center-West` + region_Southeast + 
                 capital_metrop,
               family = binomial(link = 'logit'),
               data = data_fake_news_dem)

summary(model_3)

```

Model 4 - Demographics + Political Orientation + City and Region + Religion


```{r}

model_4 <- glm(shared_fake_news ~ sex_men + age_full + race_is_white + 
                education_high + income_low + class_c + pol_orientation_right +
                pol_orientation_center + pol_orientation_left + region_North +
                region_Northeast + `region_Center-West` + region_Southeast + 
                 capital_metrop + religion_Catholic + religion_Evangelicals + 
                 `religion_Other religion`,
               family = binomial(link = 'logit'),
               data = data_fake_news_dem)

summary(model_4)

```

```{r, results = 'asis', message = FALSE}

stargazer(model_1, model_2, model_3, model_4,
          title = "Logit Models Comparison", 
          type = "latex", 
          digits = 3, 
          no.space = TRUE, 
          model.numbers = FALSE,
          header = FALSE,
          column.sep.width = "-15pt")

```





balance_table <- data_fake_news_dem %>%
  select(29:113) %>%
  lapply(., function(i) tidy(t.test(i ~ data_fake_news_dem$shared_fake_news))) %>%
  do.call(rbind, .) %>%
  rownames_to_column("variable") %>%
  rename(mean_diff = estimate, mean_control = estimate1, 
         mean_treatment = estimate2) %>%
  select(variable, mean_diff, mean_control, mean_treatment, statistic, p.value)

kable(balance_table, 
      caption = "Balance Table - Observable Characteristics",
      digits = 3,
      align = "c")


colnames(data_fake_news_dem)
































