---
title: "Fake News Data - Stats and Correlations"
author: "Natália Tosi"
date: "08/13/2021"
output: pdf_document
---

```{r message = FALSE, warning = FALSE, include = FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 8, 
  fig.height = 4
  )

options(scipen = 999)

library(tidyverse)
library(broom) 
library(stargazer)
library(knitr)
library(janitor)
library(corrplot)
library(stringr)
library(car)

```

\bigskip

```{r}

data_code <- read_csv("fake_news_data_code.csv") %>%
  select(-1)

```


# Correlation with Dependent Variable

\bigskip

**Dependent Variable = Shared Fake News (`shared_fake_news_19`)** 

\bigskip

P19 - Have you ever shared a political news story online that you thought at the time was made up? (Single Answer)

\bigskip

1 = Yes; 0 = No

\bigskip

PS: Statistically significant regarding each category to check if there are differences in observable characteristics between those who shared fake news. For example, is there a statistically significant difference between men than shared fake news and non-men that shared fake news? Analysis together with the correlation between each dummy variable with the dependent variable.

```{r}

table_correlations <- data_code %>%
  select(-P19, -shared_fake_news_19) %>%
  select_if(is.numeric) %>%
  lapply(., function(i) tidy(cor(i, data_code$shared_fake_news_19))) %>%
  do.call(rbind, .) %>%
  rownames_to_column("variable") %>%
  rename("Correlation" = x)

table_balance <- data_code %>%
  select(-P19, -shared_fake_news_19) %>%
  select(is.numeric) %>%
  lapply(., function(i) tidy(t.test(i ~ data_code$shared_fake_news_19))) %>%
  do.call(rbind, .) %>%
  rownames_to_column("variable") %>%
  rename(mean_diff = estimate, mean_control = estimate1, 
         mean_treatment = estimate2) %>%
  mutate(stat_05 = if_else(p.value < 0.05, "Yes", "No")) %>%
  select(variable, mean_diff, stat_05)

table_correlations <- table_correlations %>%
  left_join(table_balance, by = c("variable" = "variable")) %>%
  arrange(desc(Correlation))

table_correlations %>%
  kable(caption = "Balance Table - Observable Characteristics with Dependent Variable",
      digits = 3,
      align = "c")

```

\bigskip

# Multicollinearity Between Variables

\bigskip

Correlation matrix to identify multicollinearity. Excluded correlations between same variables, or dummies for the same questions (perfect collinearity addressed by dropping a value in the regression).

Highlighted variables with correlation higher than 0.3 in absolute value.

\bigskip

```{r}

data_code_numeric <- data_code %>%
  select_if(is.numeric) %>%
  select(-idInterview, -reaction_fake_news_Unsure, -reaction_fake_news_NA)
  
matrix_correlations <- as.data.frame(round(cor(data_code_numeric), 2)) %>%
  rownames_to_column("variable_1") %>%
  pivot_longer(cols = is.numeric,
               names_to = "variable_2", 
               values_to = "Correlation") %>%
  mutate(absolute_correlation = abs(Correlation)) %>%
  arrange(desc(absolute_correlation)) %>%
  mutate(question_1 = substr(variable_1, 1, 7),
         question_2 = substr(variable_2, 1, 7)) %>%
  separate(question_1, into = c("question_1", "left1"), sep = "_") %>%
  separate(question_2, into = c("question_2", "left2"), sep = "_") %>%
  mutate(same = if_else(((variable_1 == variable_2) | 
                          (question_1 == question_2)), 1, 0)) %>%
  filter(same == 0 & (absolute_correlation > 0.3)) %>%
  select(-same, -left1, -left2, -question_1, -question_2)

matrix_correlations %>%
  kable(caption = "Correlation Between Variables (> 0.3 in absolute value)",
      align = "c")

```

```{r}

matrix_correlations %>%
  ggplot(aes(x = fct_reorder(variable_1, absolute_correlation), 
             y = variable_2, fill = absolute_correlation)) + 
  geom_tile() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1, size = 6),
        axis.text.y = element_text(vjust = 1, hjust = 1, size = 6),
        plot.title = element_text(color = "grey13", size = 10, face = "bold")) +
  labs(x = "Variable 1", y = "Variable 2", fill = "Absolute Correlation", 
       title = "Correlation Between Variables (> 0.3 in absolute value)")

```



# Regressions - All categories


\bigskip

Dependent variable: `shared_fake_news_19`

\bigskip

Model 1 - Only demographics

\bigskip

```{r}

model_19_1 <- glm(shared_fake_news_19 ~ sex_men + age_full + race_is_white +
                    education_high + class_ab + class_c + has_religion, 
                  family = binomial(link = 'logit'),
                  data = data_code)

#Left out: No religion, women    

```

\bigskip

Model 2 - Demographics + Region and City

\bigskip

```{r}

model_19_2 <- glm(shared_fake_news_19 ~ sex_men + age_full + race_is_white +
                    education_high + class_ab + class_c + has_religion + 
                    region_North + region_Northeast +
                    region_Southeast + region_South + capital_metrop, 
                  family = binomial(link = 'logit'),
                  data = data_code)

```

\bigskip

Model 3 - Demographics + Region and City + Political Orientation

\bigskip

```{r}

model_19_3 <- glm(shared_fake_news_19 ~ sex_men + age_full + race_is_white +
                    education_high + class_ab + class_c + has_religion +
                    region_North + region_Northeast +
                    region_Southeast + region_South + capital_metrop + 
                    pol_orientation_right + pol_orientation_center +
                    pol_orientation_left,
                  family = binomial(link = 'logit'),
                  data = data_code)

#Left out: religion_No religion + region_Center-West + pol_orientation_none                             

```

\bigskip

Model 4 - Demographics + Region and City + Political Orientation + Government Evaluation

\bigskip

```{r}

model_19_4 <- glm(shared_fake_news_19 ~ sex_men + age_full + race_is_white +
                    education_high + class_ab + class_c + has_religion + 
                    region_North + region_Northeast +
                    region_Southeast + region_South + capital_metrop + 
                    pol_orientation_right + pol_orientation_center +
                    pol_orientation_left  + approves_gov,
                  family = binomial(link = 'logit'),
                  data = data_code)

#Left out: religion_No religion + region_Center-West + pol_orientation_none            

```


\bigskip

Model 5 - Demographics + Region and City + Political Orientation + Government Evaluation + Answers Fake News

\bigskip

```{r}

model_19_5 <- glm(shared_fake_news_19 ~ sex_men + age_full + race_is_white +
                    education_high + class_ab + class_c + has_religion + 
                    region_North + region_Northeast +
                    region_Southeast + region_South + capital_metrop + 
                    pol_orientation_right + pol_orientation_center +
                    pol_orientation_left + approves_gov + 
                    frequency_fake_news + resp_population + resp_gov + 
                    resp_politicians + resp_press + resp_social_media +
                    severity_fake_news + fact_checking,
                  family = binomial(link = 'logit'),
                  data = data_code)

```

\bigskip

Model 6 - Demographics + Region and City + Political Orientation + Government Evaluation + Pandemic

\bigskip

*attention increased interest in science* 

```{r}

model_19_6 <- glm(shared_fake_news_19 ~ sex_men + age_full + race_is_white +
                    education_high + class_ab + class_c + has_religion + 
                    region_North + region_Northeast +
                    region_Southeast + region_South + capital_metrop + 
                    pol_orientation_right + pol_orientation_center +
                    pol_orientation_left  + approves_gov + 
                    pand2_worse_perception_media + pand3_trust_vaccine + 
                    pand3_seek_science + pand3_preventive_treat + pand3_masks +
                    pand5_increased_interest_science, 
                  family = binomial(link = 'logit'),
                  data = data_code)

```

\bigskip

Model 7 - Demographics + Region and City + Political Orientation + Government Evaluation + Vote

\bigskip

```{r}

model_19_7 <- glm(shared_fake_news_19 ~ sex_men + age_full + race_is_white +
                    education_high + class_ab + class_c + has_religion + 
                    region_North + region_Northeast +
                    region_Southeast + region_South + capital_metrop + 
                    pol_orientation_right + pol_orientation_center +
                    pol_orientation_left + approves_gov + 
                    vote1_trust_ballot + vote2_eletronic_best_option + 
                    vote3_worried_hacker + vote3_worried_politics +
                    vote3_worried_transparency + vote3_worried_tech + 
                    vote3_worried_tse,
                  family = binomial(link = 'logit'),
                  data = data_code)

#Left out: religion_Other religion + pol_orientation_none    

```

\bigskip

Model 8 - All

\bigskip

```{r}

model_19_8 <- glm(shared_fake_news_19 ~ sex_men + age_full + race_is_white +
                    education_high + class_ab + class_c + has_religion + 
                    region_North + region_Northeast +
                    region_Southeast + region_South + capital_metrop + 
                    pol_orientation_right + pol_orientation_center +
                    pol_orientation_left + approves_gov +
                    frequency_fake_news + resp_population + resp_gov + 
                    resp_politicians + resp_press + resp_social_media +
                    severity_fake_news + fact_checking + 
                    pand2_worse_perception_media + pand3_trust_vaccine + 
                    pand3_seek_science + pand3_preventive_treat + pand3_masks +
                    pand5_increased_interest_science + vote1_trust_ballot + 
                    vote2_eletronic_best_option + vote3_worried_hacker + 
                    vote3_worried_politics + vote3_worried_transparency + 
                    vote3_worried_tech + vote3_worried_tse,
                  family = binomial(link = 'logit'),
                  data = data_code)

```

\bigskip


# Models

```{r, results = 'asis', message = FALSE, echo = FALSE}

stargazer(model_19_1, model_19_2, model_19_3, model_19_4,
          title = "Logit Models Comparison - Up to Government Evaluation", 
          type = "latex", 
          digits = 3, 
          no.space = TRUE, 
          model.numbers = FALSE,
          header = FALSE,
          font.size = "footnotesize",
          column.sep.width = "-15pt")

```

\newpage

```{r, results = 'asis', message = FALSE, echo = FALSE}

stargazer(model_19_1, model_19_2, model_19_3, model_19_5,
          title = "Logit Models Comparison - Answers Fake News", 
          type = "latex", 
          digits = 3, 
          no.space = TRUE, 
          model.numbers = FALSE,
          header = FALSE,
          font.size = "footnotesize",
          column.sep.width = "-15pt")

```

\newpage

```{r, results = 'asis', message = FALSE, echo = FALSE}

stargazer(model_19_1, model_19_2, model_19_3, model_19_6,
          title = "Logit Models Comparison - Pandemic", 
          type = "latex", 
          digits = 3, 
          no.space = TRUE, 
          model.numbers = FALSE,
          header = FALSE,
          font.size = "footnotesize",
          column.sep.width = "-15pt")

```

\newpage

```{r, results = 'asis', message = FALSE, echo = FALSE}

stargazer(model_19_1, model_19_2, model_19_3, model_19_7,
          title = "Logit Models Comparison - Electronic Vote", 
          type = "latex", 
          digits = 3, 
          no.space = TRUE, 
          model.numbers = FALSE,
          header = FALSE,
          font.size = "footnotesize",
          column.sep.width = "-15pt")

```

\newpage

```{r, results = 'asis', message = FALSE, echo = FALSE}

stargazer(model_19_5, model_19_6, model_19_7, model_19_8,
          title = "Logit Models Comparison - All Variables", 
          type = "latex", 
          digits = 3, 
          no.space = TRUE, 
          model.numbers = FALSE,
          header = FALSE,
          font.size = "tiny",
          column.sep.width = "-15pt")

```

\newpage

# Variance Inflation Factor (VIF)

\bigskip

"For a given predictor (p), multicollinearity can assessed by computing a score called the variance inflation factor (or VIF), which measures how much the variance of a regression coefficient is inflated due to multicollinearity in the model.

The smallest possible value of VIF is one (absence of multicollinearity). As a rule of thumb, a VIF value that exceeds 5 or 10 indicates a problematic amount of collinearity (James et al. 2014)."

```{r}

model_1 <- as.data.frame(car::vif(model_19_1)) %>%
  rownames_to_column("variables")

model_2 <- as.data.frame(car::vif(model_19_2)) %>%
  rownames_to_column("variables")

model_3 <- as.data.frame(car::vif(model_19_3)) %>%
  rownames_to_column("variables")

model_4 <- as.data.frame(car::vif(model_19_4)) %>%
  rownames_to_column("variables")

model_5 <- as.data.frame(car::vif(model_19_5)) %>%
  rownames_to_column("variables")

model_6 <- as.data.frame(car::vif(model_19_6)) %>%
  rownames_to_column("variables")

model_7 <- as.data.frame(car::vif(model_19_7)) %>%
  rownames_to_column("variables")

model_8 <- as.data.frame(car::vif(model_19_8)) %>%
  rownames_to_column("variables")

vif_test <- model_8 %>%
  left_join(model_1, by = c("variables" = "variables")) %>%
  left_join(model_2, by = c("variables" = "variables")) %>%
  left_join(model_3, by = c("variables" = "variables")) %>%
  left_join(model_4, by = c("variables" = "variables")) %>%
  left_join(model_5, by = c("variables" = "variables")) %>%
  left_join(model_6, by = c("variables" = "variables")) %>%
  left_join(model_7, by = c("variables" = "variables"))

names(vif_test) <- c("variables", "model_8", "model_1", "model_2", "model_3",
                     "model_4", "model_5", "model_6", "model_7")

vif_test <- relocate(vif_test, -model_8)

vif_test %>%
  kable(caption = "Variance Inflation Factor (VIF) per variable and model",
      align = "c")


```

